---
title: "power_rmd"
author: "DVM Bishop"
date: "01/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(umx)
```

Double checking power calculations, after query by Guy Vingerhoets.

Using script from https://cran.r-project.org/web/packages/umx/index.html, see p 23.

Ideally, I'd like to make a table from the output, but it is frustratingly hard to find out how to access the values that are created, so it's currently just a long dump into a text file.

NOTE: the value of AA entered into this function is proportion of variance explained by genetic factors in AE model (after dropping c), i.e equivalent to standardized a^2.

```{r powerACE}
myAA<-c(.1,.15,.2,.25,.3,.35) #values of a2 to test
twins1<-c(96,98) #N MZ/DZ twin pairs, EHI
twins2<-c(65,76) #N twin pairs, LI


for (i in 1:6){
  for (mypower in seq(.6,.9,.05)){
  print(paste0('power = ',mypower))
  thisAA<-myAA[i]
   print(paste0('AA = ',thisAA))

thispower<-power.ACE.test(
AA = thisAA,
CC = 0,
EE = NULL,
update = c("a_after_dropping_c"),
value = 0,
n = NULL,
MZ_DZ_ratio = 1,
sig.level = 0.05,
power =mypower,
method = c("ncp"),
search = FALSE,
tryHard = c("yes"),
optimizer = NULL,
nSim = 4000
)
  }
}

```